---
title: "x-field_sandbox"
author: "Elizabeth Elton"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(hablar)
library(ggbeeswarm)
library(plotly)
library(FactoMineR)
library(factoextra)
cwd <- here::here()

library(fields)
library(bigsplines)
#library(plot3D)

source('chip_invasion/functions.R', local = TRUE)
```

#read in data using normal functions
#find xmin and xmax
#divide into fields
#label xy fields -- change to letters? or it can be like 1_1, 1_2, etc
#change endo boundary calc to be grouped by xy fields
#get counts
#compare to ground truth -- ask carly/rachel for this

#hct data -- read in using normal fxns
#get colnames/order
#build in rearranger to make colnames the order i want

```{r}
diagnostic_plots_ee <- function(huvec_df, organoid_df) {
  huvec_plot <- ggplot(huvec_df, aes(x = centroid_z)) +
    geom_histogram() +
    facet_wrap(~ chip) +
    geom_vline(xintercept = mean(huvec_df$centroid_z))+
    labs(title = "Endothelial cells at each z position")
  
  organoid_plot <- ggplot(organoid_df, aes(x = centroid_z))  +
    geom_histogram() +
    facet_wrap(~ chip) +
    labs(title = "Epithelial cells at each z position")
  
  print(huvec_plot)
  print(organoid_plot)
}
```



```{r}
test_us <- endothelial_epithelial_reader("/Users/eelton/Dropbox (EITM)/CLS Files Share with Elizabeth 2023/US-GFP Organoids CLS Analysis/20230219exp_D6_Chip1-6_invasion_US-GFP ORGS.xlsx")
diagnostic_plots_ee(huvec_df = test_us$huvec_df, test_us$organoid_df)
```


```{r}
meantest <- test_us$huvec_df %>%
  group_by(chip) %>%
  summarise(xint = mean(centroid_z))

huvec_df <- test_us$huvec_df

ggplot(huvec_df, aes(x=position_x, y=centroid_z)) +
  geom_point() +
  facet_wrap(~ chip)

#lets start w one chip
chip1 <- huvec_df %>%
  filter(chip == 1)

ggplot(chip1, aes(x=position_x, y=position_y)) +
  geom_point(aes(color=centroid_z)) +
  geom_vline(xintercept = widthlist$xfield_bounds)
#cool field assignment is working

ggplot(chip1, aes(x=position_x)) +
  geom_histogram(binwidth = 1)
```


```{r}
chip1_xmin <- min(chip1$position_x)
chip1_xmax <- max(chip1$position_x)
chip1_xfield_size <- (chip1_xmax-chip1_xmin)/10
widthlist <- list()
for (f in 0:10){
  xfieldbound <- chip1_xmax - (f*chip1_xfield_size)
  
  widthlist <- append(widthlist, xfieldbound)
} #these are the xfield boundaries

widthlist <- as.data.frame(widthlist) %>%
  pivot_longer(cols = 1:11, names_to = "idc", values_to = "xfield_bounds")

chip1 <- chip1 %>%
  mutate(xfield = case_when(widthlist$xfield_bounds[2] <= position_x ~ 10,
                            widthlist$xfield_bounds[3] <= position_x & position_x <= widthlist$xfield_bounds[2] ~ 9,
                            widthlist$xfield_bounds[4] <= position_x & position_x <= widthlist$xfield_bounds[3] ~ 8,
                            widthlist$xfield_bounds[5] <= position_x & position_x <= widthlist$xfield_bounds[4] ~ 7,
                            widthlist$xfield_bounds[6] <= position_x & position_x <= widthlist$xfield_bounds[5] ~ 6,
                            widthlist$xfield_bounds[7] <= position_x & position_x <= widthlist$xfield_bounds[6] ~ 5,
                            widthlist$xfield_bounds[8] <= position_x & position_x <= widthlist$xfield_bounds[7] ~ 4,
                            widthlist$xfield_bounds[9] <= position_x & position_x <= widthlist$xfield_bounds[8] ~ 3,
                            widthlist$xfield_bounds[10] <= position_x & position_x <= widthlist$xfield_bounds[9] ~ 2,
                            position_x <= widthlist$xfield_bounds[10] ~ 1),
         xfield = as.character(xfield),
         yfield = as.character(field))
#there is for sure a better way to do this  
#cool now i have field assign,ent, now 

```

```{r}
#trying tps!
#silly test - trying a 2d example. i am concerned by the time this is taking
#fit <- fastTps(chip1$position_x, chip1$position_y, aRange = 3)

fit <- fastTps(x=chip1[, c("position_x", "position_y")], Y=chip1$centroid_z, aRange = 200)
fit_100 <- fastTps(x=chip1[, c("position_x", "position_y")], Y=chip1$centroid_z, aRange = 100)

#fit_centroid <- fastTps(x=chip1[, c("centroid_x_in_image", "centroid_y_in_image")], Y=chip1$centroid_z, aRange = 200)
#set.panel(2,2)
plot(fit)
#set.panel()

summary(fit)

surface(fit)
#!!!
#stick with fasttps for now, see if i can get it to work w predict
summary(fit_100)
surface(fit_100)

```
```{r}
surface(fit)
```



```{r}
#try predicsurface from fields -- use grid of values from??

#fit.pred <- predictSurface(fit)
#summary(fit.pred)
fit.pred <- predict(fit)
summary(fit.pred)
#surface(fit)
```

```{r}
quilt.plot(x=chip1$position_x, y=chip1$position_y, z=chip1$centroid_z)
#oh this is kind of a cool visual
```
```{r}
grid.list <- list(x=seq(min(chip1$position_x), max(chip1$position_x)), y=seq(min(chip1$position_y), max(chip1$position_y)))
grid.list <- list(x=seq(1,65), y=seq(1,66.04615))
dat <- as.matrix(chip1$centroid_z, nr=65, nc=65)

full.grid <- make.surface.grid(grid.list = grid.list)
new.fit <- spatialProcess(x=chip1[, c("position_x", "position_y")], y=chip1$centroid_z) #takes forever dont do this

summary(new.fit)
```


```{r}
surface(new.fit) #that looks so cool but took so long
```


```{r}
xa.2 <- seq(0, max(chip1$position_x), len = 100) #xposition -- of org data?? what data do i want here -- start w huvec just to see
xb.2 <- seq(0, max(chip1$position_y), len = 100) #yposition
data250 <- expand.grid(x=xa.2, y=xb.2) #make big grid -- try field built in?
fitted.add1215 <- matrix(predict(fit, data250), 100, 100) #compare fit to grid

#grid.list <- list(x=seq(0, max(chip1$position_x), len = 1000), y=seq(0, max(chip1$position_y), len = 1000))
#full.grid <- make.surface.grid(grid.list = grid.list)

#fitted.add1215 <- matrix(predict(fit, full.grid), 1000, 1000)
#ok dont use their make.surface.grid fxn
#3d plot
#open3d()
# material3d(col="black")
#persp3d(z = fitted.add1215, contour = TRUE, theta = 70, phi=30, ticktype="detailed", xlab=drugA, ylab=drugB,  zlab="Viability", nlevels = 3, col= viridis(1000), main = 'Simulated Additive LIM1215')

fig <- plot_ly(z = fitted.add1215) %>% add_surface()

fig
```


```{r}
chip1_gfps <- test_us$organoid_df %>%
  filter(chip==1) %>%
  select(position_x, position_y)

#make a list by chip 
#run lappy
#unlist and cbind (rbind and cbind)

#predict(model, (xcords, ycords))
gfp_pred <- predict(fit, chip1_gfps)
#i mean cool
#so now we try it as mutate and compare

chip_gfp_comp <- test_us$organoid_df %>%
  filter(chip==1)
  #mutate(predicted_z = predict(fit, c(position_x, position_y)),
   #      below_pred = centroid_z < predicted_z)
#oh i dont think i can mutate

chip_gfp_comp <- cbind(chip_gfp_comp, gfp_pred) %>%
  mutate(below_pred = centroid_z < gfp_pred)
#sure
chip_gfp_bottom <- chip_gfp_comp %>%
  filter(below_pred == TRUE)
#that seems like a lot #because it is all chips!!! not just chip 1!!! im silly
#ok but i havent filtered by intensity, pca yet

ggplot(chip_gfp_bottom) +
  geom_point(aes(x=position_x, y=centroid_z))
#oh cool!!! 38 seems reasonable maybe
#see what it looks like when i use the app
#and when i run intensity and pca filters on this

plot_ly(x=chip_gfp_bottom$position_x, y=chip_gfp_bottom$position_y, z=chip_gfp_bottom$centroid_z)
```

```{r}
ggplot(chip_gfp_bottom, aes(x = position_x, y = position_y)) + geom_point() + labs(title = "XY Position Plot")
ggplot(chip_gfp_bottom, aes(x=position_x, y=centroid_z)) + geom_point() + labs(title = "XZ Position Plot")
```



```{r}
intensity_test <- org_intensity_filtering(chip_gfp_bottom)
high_int_test <- intensity_test$gfp_high_intensity #down to 19

#cleanpcatest <- ee_invasion_pcar(intensity_test$gfp_high_intensity) #hmmmm figure this out later !!!!!

cleanpcatest  <- intensity_test$gfp_high_intensity %>%
    select(-(5:8), -(11:17), -(19:27), -(37:39), -(44:46)) %>% #select relevant cols
    mutate(chip = as_factor(chip), field = as_factor(field)) 

#pcamatrixtest <- ee_invasion_pcar(cleanpcatest)

nointensity.pca <- cleanpcatest %>%
    mutate_if(is.numeric, scale) %>%
    PCA(quanti.sup = c(1:3, 5:7, 19), quali.sup = c(4, 20:21), graph = FALSE)

invasion_pcar_graphs(nointensity.pca, cleanpcatest)

pcafilttest <- pcar_density_filter(cleanpcatest, nointensity.pca)
pcafilttest$pca_filtered_df #still 19


pcaclean <- ee_clean_for_pca(intfilt$gfp_high_intensity)

pcamatrix <- ee_invasion_pcar(pcaclean)

invasion_pcar_graphs(pcamatrix, pcaclean)

pcafiltered <- pcar_density_filter(pcaclean, pcamatrix)

per_chip_counter_ee(pcafiltered$pca_filtered_df)
```



```{r}
# fit_better <- bigtps(x=chip1[, c("position_x", "position_y")], y=chip1$centroid_z)
# summary(fit_better)
# imagebar(x=fit_better$x[,1], y=fit_better$x[,2], z=fit_better$fitted.values)
# surface(fit_better)
#persp3d or surf 3d
#x=arranged xinput
#y
```

```{r}
#runing thru the normal pipeline

step1 <- endothelial_epithelial_reader("/Users/eelton/Dropbox (EITM)/CLS Files Share with Elizabeth 2023/US-GFP Organoids CLS Analysis/20230219exp_D6_Chip1-6_invasion_US-GFP ORGS.xlsx")
endobound <- endo_boundary(step1$huvec_df, step1$cell_type)
firstcounts <- epi_count(step1$organoid_df, endobound$field_chip_huvec)
firstcounts$gfp_per_chip_bottom #47 pre int,pca #hmm i would have expected this to be lower

chip_3d_plot_ee(firstcounts$all_gfp_bottom, chip_in = 1)
position_chip_plots_xy(firstcounts$all_gfp_bottom)
position_chip_plots_xz(firstcounts$all_gfp_bottom)
```


```{r}
intfilt <-org_intensity_filtering(firstcounts$all_gfp_bottom)

intfilt_c1 <- intfilt$gfp_high_intensity %>% filter(chip == 1)
per_chip_counter_ee(intfilt$gfp_high_intensity) #23

pcaclean <- ee_clean_for_pca(intfilt$gfp_high_intensity)

pcamatrix <- ee_invasion_pcar(pcaclean)

invasion_pcar_graphs(pcamatrix, pcaclean)

pcafiltered <- pcar_density_filter(pcaclean, pcamatrix)

per_chip_counter_ee(pcafiltered$pca_filtered_df) #12
```

#trying on several chips
```{r}
huveclist <- split(test_us$huvec_df, f=test_us$huvec_df$chip)

fit_multi <- lapply(huveclist, function(x) fastTps(x[, c("position_x", "position_y")], Y=x$centroid_z, aRange = 200))
names(fit_multi) <- c("chip1", "chip2", "chip3", "chip4", "chip5", "chip6")
#took like 30sec

orglist <- split(test_us$organoid_df, f=test_us$organoid_df$chip)
names(orglist) <- c("chip1", "chip2", "chip3", "chip4", "chip5", "chip6")

#org_positions <- lapply(orglist, function(x) dplyr::select(x, position_x, position_y))

#org_predictions <- lapply(org_positions, function(x) predict(fit_multi, x))

#org_predictions <- lapply(org_positions, function(x) lapply(fit_multi, function(y) predict(y, x)))
#i dont think this is correct
chip1pred <- as.data.frame(org_predictions$`1`[1])
gfp_pred <- as.data.frame(gfp_pred) %>% rename(X1=V1)
#these are the same

# testlist <- list()
# for (i in names(orglist)) {
#   positions <- as_data_frame(c(orglist$i$position_x),
#                              c(orglist$i$position_y))
#     #dplyr::select(orglist$`i`, position_x, position_y)
#   
#   modelly <- fit_multi$i
#   
#   predicitions <- predict(modelly, positions)
#   
#   df_w_predictions <- cbind(orglist$i, predicitions)
#   
#  testlist <- append(testlist, df_w_predictions) 
# }

org_positions <- lapply(orglist, function(x) dplyr::select(x, position_x, position_y))



#oh wait
hmmmmmm <- map2(.x=fit_multi, .y=org_positions, ~predict(.x, .y))
#...did that work?
#oh shit it mightve
hmx2 <- map2(.x=orglist, .y=hmmmmmm, ~cbind(.x, .y))
hmx2clean <- lapply(hmx2, function(x) rename(x, "gfp_pred"=.y))

#oooooh we're doing it now
preds_together <- bind_rows(hmx2clean)
#nice!
#ok so now each chip has its predictions, i need to mutate to just get those below the pred and then count per chip
allchips <- preds_together %>%
  group_by(chip) %>%
  mutate(below_pred = centroid_z < gfp_pred) %>%
  filter(below_pred == TRUE)

countsz <- per_chip_counter_ee(allchips)
countsz

position_chip_plots_xy(allchips)
position_chip_plots_xy(allchips)
chip_3d_plot_ee(allchips, 6)
#oh cool!!! 38 seems reasonable maybe
#see what it looks like when i use the app
#and when i run intensity and pca filters on this

plot_ly(x=chip_gfp_bottom$position_x, y=chip_gfp_bottom$position_y, z=chip_gfp_bottom$centroid_z)
```


```{r}
int_allchips <- org_intensity_filtering(allchips)
per_chip_counter_ee(int_allchips$gfp_high_intensity)
```


```{r}
surface(fit_multi$chip2)
```

```{r}
x2 <- seq(0, max(huveclist$`2`$position_x), len = 100)  #this isnt what i want
y2 <- seq(0, max(huveclist$`2`$position_y), len = 100) #yposition
data100 <- expand.grid(x=x2, y=y2) #make big grid -- try field built in?
fitted.chip2 <- matrix(predict(fit_multi$chip2, data100), 100, 100) #compare fit to grid

fig <- plot_ly(z = fitted.chip2) %>% add_surface()

fig #that looks crazy!!!!! need to ask them if this is a real thing they are seeing
```


```{r}
endo_boundary_test <- function(huvec_df, cell_type) {
  
  mean_huvec_pop <- huvec_df %>% 
    # filter(`546 Upper Spot Bright - 546 Spot Bright Image Region Centroid Z [µm]` >= channel_boundary_number )  %>% 
    group_by(chip) %>% 
    summarise(mean_top_bottom = mean(centroid_z), 
              median_top_bottom = median(centroid_z),
              sd_top_bottom = sd(centroid_z)
    ) %>% 
    mutate(lower_cuttoff = mean_top_bottom - (2 * sd_top_bottom),
           upper_cuttoff = mean_top_bottom + (2 * sd_top_bottom))
  
  middle_chip_huvec <- huvec_df %>% 
    inner_join(mean_huvec_pop) %>%
    group_by(chip, field) %>%
    mutate(top_bottom_2sd_ll = centroid_z >= lower_cuttoff,
           top_bottom_2sd_ul = centroid_z <= upper_cuttoff)
  
  if(cell_type == "organoid"){
    middle_chip_huvec <- middle_chip_huvec %>%
      filter(
        # Field %in% c(4, 5, 6, 7, 8, 9, 10, 11, 12),
        top_bottom_2sd_ll == TRUE, 
        top_bottom_2sd_ul == TRUE)
  }
  
  field_chip_huvec <- middle_chip_huvec %>%
    dplyr::summarise(mean_huvec_z_per_field = mean(centroid_z), 
                     sd_huvec_z_per_field = sd(centroid_z),
                     huvec_count = n()) %>% 
    mutate(mean_huvec_offset = case_when(cell_type == "hct" ~ (mean_huvec_z_per_field + (0.7*(mean_huvec_z_per_field / sd_huvec_z_per_field))),
                                         cell_type == "organoid" ~ (mean_huvec_z_per_field - (0.25*sd_huvec_z_per_field)))) %>%
    mutate(mean_huvec_offset_old = (mean_huvec_z_per_field + (1.1*(mean_huvec_z_per_field / sd_huvec_z_per_field)))) %>%
    ungroup()
  
  field_chip_huvec_summary <- field_chip_huvec %>% 
    dplyr::select(chip, field, mean_huvec_z_per_field, mean_huvec_offset) %>%
    mutate(mean_huvec_z_per_field = round(mean_huvec_z_per_field, 2),
           mean_huvec_offset = round(mean_huvec_offset, 2)) %>%
    dplyr::rename("Chip" = chip, "Field" = field, "Mean Endothelial Z-height" = mean_huvec_z_per_field, "Endothelial Cutoff Z-height" = mean_huvec_offset)
  
  huvec_dfs <- list("field_chip_huvec" = field_chip_huvec, "field_chip_huvec_summary" = field_chip_huvec_summary)
  
  return(huvec_dfs)
}
```


